`r glue::glue('inferCNV results for {params$library_id} with {params$reference_name} reference')`

### 2025-05-07 analyses/infercnv-consensus-cell-type/template-notebooks/SCPCP000015_explore-infercnv-results.Rmd
cluster plots

```{r, fig.height = 10}
# Filter to just cells in the library to avoid warnings
query_has_cnv_df <- has_cnv_df |>
  dplyr::filter(library_id == params$library_id)

celltype_umap <- ggplot(query_has_cnv_df) +
  aes(x = UMAP1, y = UMAP2, color = celltype_n) +
  geom_point(size = 0.1, alpha = 0.3) +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1))) +
  labs(title = "Cell types") +
  umap_theme

cnv_umap <- ggplot(query_has_cnv_df) +
  aes(x = UMAP1, y = UMAP2, color = total_cnv_per_cell) +
  geom_point(size = 0.1, alpha = 0.3) +
  scale_color_viridis_c() +
  labs(title = "Total CNV per cell") +
  umap_theme

@@

```{r}
cluster_count_df <- has_cnv_df |>
  dplyr::count(subcluster, cell_group)
ggplot(cluster_count_df) +
  aes(x = n, fill = cell_group) +
  geom_histogram(color = "grey20", bins = 15) +
  facet_wrap(vars(cell_group), scales = "free")
```

@@

Let's also look more specifically at the `reference` clusters, where we might expect some batch effects since cells were taken from several different libraries.
Do originating libraries cluster together?
When interpreting this plot, bear in mind that there might also be an association between cell types and libraries, and we would expect cell types to still cluster together.

>>>>

```{r, fig.height = 10}
# Filter to just cells in the library to avoid warnings
library_has_cnv_df <- has_cnv_df |>
  dplyr::filter(library_id == params$library_id)

celltype_umap <- ggplot(library_has_cnv_df) +
  aes(x = UMAP1, y = UMAP2, color = celltype_n) +
  geom_point(size = 0.1, alpha = 0.3) +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1))) +
  labs(title = "Cell types") +
  umap_theme

cnv_umap <- ggplot(library_has_cnv_df) +
  aes(x = UMAP1, y = UMAP2, color = total_cnv_per_cell) +
  geom_point(size = 0.1, alpha = 0.3) +
  scale_color_viridis_c() +
  labs(title = "Total CNV per cell") +
  umap_theme

@@

```{r}
cluster_count_df <- has_cnv_df |>
  dplyr::count(subcluster, cell_group_n)
ggplot(cluster_count_df) +
  aes(x = n, fill = cell_group_n) +
  geom_histogram(color = "grey20", bins = 15) +
  facet_wrap(vars(cell_group_n), scales = "free") +
  labs(
    x = "Number of cells in cluster",
    y = "Number of clusters"
  )
```

@@

We'd also like to look at the distribution of cell types across clusters, specifically considering the non-reference clusters.
For this, we'll focus on the top 10 cell types (based on the `cell-type-ewings` labels) considering all "tumor" as a single grouping.
For visual ease, we'll only show clusters with at least 10 cells.

```{r}
query_only_df <- library_has_cnv_df |>
  dplyr::filter(cell_group == "unknown") |>
  # create a finer cell type column for this plot
  dplyr::mutate(
    celltype_fine = dplyr::case_when(
      cell_group == "reference" ~ "reference cell",
      stringr::str_detect(ewing_annotation, "tumor") ~ "tumor cell",
      .default = ewing_annotation
    )
  ) |>
  dplyr::add_count(celltype_fine, name = "celltype_fine_count") |>
  dplyr::mutate(
    celltype_fine_n = glue::glue("{celltype_fine} (n={celltype_fine_count})"),
    # top 10 only
    celltype_fine_n = forcats::fct_lump(celltype_fine_n, 10),
  )

# add n() to "Other" for plot clarity and update factor to frequency order
current_levels <- levels(query_only_df$celltype_fine_n)
total_other <- query_only_df |>
  dplyr::filter(celltype_fine_n == "Other") |>
  nrow()
updated_other <- glue::glue("Other (n={total_other})")

query_only_df <- query_only_df |>
  dplyr::mutate(
    # update other level to have counts
    celltype_fine_n = as.character(celltype_fine_n),
    celltype_fine_n = ifelse(celltype_fine_n == "Other", updated_other, celltype_fine_n),
    # order based on freq
    celltype_fine_n = as.factor(celltype_fine_n),
    celltype_fine_n = forcats::fct_infreq(celltype_fine_n),
    celltype_fine_n = forcats::fct_relevel(celltype_fine_n, updated_other, after = Inf)
  ) |>
  # exclude small clusters
  dplyr::add_count(subcluster, name = "subcluster_size") |>
  dplyr::filter(subcluster_size >= 10)

# 11 hex colors!
# adapted from 'Classic 10' from paletteer: https://r-graph-gallery.com/color-palette-finder
colors <- c("#1F77B4", "#FF7F0E", "#2CA02C", "#D6272F", "#9467BD", "#8C564B", "#E377C2", "#7F7F7F", "#BCBD22", "#17CFFB", "#17CF1F")

ggplot(query_only_df) +
  aes(
    x = forcats::fct_infreq(subcluster),
    fill = celltype_fine_n
  ) +
  geom_bar() +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_blank())
```

Finally, let's also look more specifically at the `reference` clusters, where we might expect some batch effects since cells were taken from several different libraries.
Do originating libraries cluster together?
When interpreting this plot, bear in mind that there might also be an association between cell types and libraries, and we would expect cell types to still cluster together.
